---
title: "Fine-Gray Competing Risks Regression with mlr3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fine-Gray Competing Risks Regression with mlr3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("mlr3proba", quietly = TRUE) &&
         requireNamespace("cmprsk", quietly = TRUE) &&
         requireNamespace("LearnerCompRisksFineGrayCRR", quietly = TRUE) &&
         requireNamespace("future", quietly = TRUE) &&
         requireNamespace("future.apply", quietly = TRUE)
)
```

# Introduction

This vignette demonstrates the `cmprsk.crr` learner from
the `LearnerCompRisksFineGrayCRR` package, which implements
Fine-Gray competing risks regression using `cmprsk::crr`. 
The learner estimates cumulative incidence functions (CIFs)
for competing risks scenarios, supporting fixed and time-varying
ovariates. It is designed for tasks with multiple mutually exclusive events,
such as survival analysis with competing risks. For detailed documentation, 
see `?LearnerCompRisksFineGrayCRR::mlr_learners_cmprsk.crr`.

# Setup

Load required packages and create a task using the `pbc` dataset:

```{r setup}
library(mlr3)
library(mlr3proba)
library(LearnerCompRisksFineGrayCRR)
if (!requireNamespace("cmprsk", quietly = TRUE)) {
  stop("Package 'cmprsk' must be installed")
}
task <- tsk("pbc")
task$set_col_roles(cols = "status", add_to = "stratum")
task$col_roles$feature <- setdiff(task$feature_names, "status")
task$select(c("age", "sex", "bili"))
print(task)
print(table(task$data()[["status"]]))
set.seed(123)
part <- partition(task, ratio = 0.7)
sapply(part, length) # partition sizes

```

The `pbc` task is configured with age, sex, and bili as features and status
as the stratum for competing events.




# Training the Learner

## Fixed covariates only


Train the Fine-Gray model with fixed (time-invariant) covariates (no time-varying effects):

```{r train_fixed}
learner <- lrn("cmprsk.crr", maxiter = 100, gtol = 1e-6, parallel = FALSE)
learner$train(task, row_ids = part$train)
print(learner)
print(learner$model)
```

The learner is initialized with default parameters: 
`maxiter = 100` for maximum iterations and `gtol = 1e-6` for convergence tolerance.




## Time-Varying Covariates


`
Train the model with time-varying covariates, where `bili` is used only
as a time-varying effect:

```{r train_time_varying}
learner_tv <- lrn("cmprsk.crr",
  cov2_info = list(
    cov2nms = c("age", "bili"),
    tf = function(uft) cbind(log(uft), log(uft + 1)),
    cov2only = "bili"
  ),
  maxiter = 50,
  gtol = 1e-7,
  parallel = FALSE
)
learner_tv$train(task, row_ids = part$train)
print(learner_tv$param_set)
print(learner_tv$model)

```


The `cov2_info` parameter specifies time-varying covariates (age and bili)
with a transformation function (tf) that applies logarithmic transformations
to event times. The `cov2only = "bili"` ensures bili is used only 
as a time-varying covariate, not as a fixed effect.

## Parallel Processing

For datasets with many events, parallel processing can reduce computation time.
This requires the `future.apply` package and a future plan:

```{r train_parallel, eval = requireNamespace("future.apply", quietly = TRUE) && requireNamespace("future", quietly = TRUE)}
on.exit(future::plan("sequential"), add = TRUE)
future::plan("multisession")
learner_parallel <- lrn("cmprsk.crr", parallel = TRUE)
learner_parallel$train(task, row_ids = part$train)
print(learner_parallel$param_set)
```

Set `future::plan("multisession")` or another plan before enabling 
`parallel = TRUE`. This is particularly useful for large datasets 
or tasks with many competing events.




# Prediction

Predict CIFs for the test set:

```{r predict}
pred <- learner$predict(task, part$test)
print(pred)
class(pred)

```

The predictions are `PredictionCompRisks` objects, containing CIFs for each event as 
data tables, with rows corresponding to observations and columns to unique event
times from the training data.


```{r plot_cif1, fig.align = "center", fig.width = 7, fig.height = 5, out.width = "80%"}

# Extract cif1
cif1 <- pred$cif[["1"]]  # Matrix: 82 x 103
dim(cif1)  # Confirm: [1] 82 103

# Plot CIF curves for 5 randomly selected observations
set.seed(123)
sample_rows <- sample(1:82, 5)
time_points <- as.numeric(colnames(cif1))  # Extract numeric time points
matplot(
  time_points,
  t(cif1[sample_rows, ]),
  type = "l",
  lty = 1,
  col = 1:5,
  xlab = "Time",
  ylab = "Cumulative Incidence",
  main = "Cumulative Incidence Functions for Event 1",
  ylim = c(0, 1)
)
legend(
  "topleft",
  legend = paste("Obs", task$row_names[part$test][sample_rows]),
  col = 1:5,
  lty = 1
)
```

This plot shows the cumulative incidence curves for event 1 
for five randomly selected test observations, 
with time on the x-axis and CIF values (0 to 1) on the y-axis.



# Model Diagnostics

Check convergence and variable importance:

```{r diagnostics}
learner$convergence()
learner$importance()
learner_tv$convergence()
learner_tv$importance()

```

* *Convergence:* The `convergence()` method returns a named list with TRUE or FALSE 
for each event, indicating whether the model converged. Non-convergence may occur
with strict gtol (e.g., 1e-9) or low maxiter (e.g., 1).

* *Importance:* The `importance()` method returns a data frame with normalized absolute coefficients, 
showing the relative measure of each variable influence for each event. Higher values indicate greater
importance.

## See Also

* `?LearnerCompRisksFineGrayCRR::mlr_learners_cmprsk.crr` for learner documentation.
* `?cmprsk::crr` for details on the underlying Fine-Gray model.
* `?mlr3proba::LearnerCompRisks` for the base class used by this learner


```{r cleanup, include = FALSE}
future::plan("sequential")
```




References

Fine, J. P., & Gray, R. J. (1999). A Proportional Hazards Model for the Subdistribution of a Competing Risk. Journal of the American Statistical Association, 94(446), 496--509. doi:10.1080/01621459.1999.10474144







