---
title: "Fine-Gray Competing Risks Regression with mlr3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fine-Gray Competing Risks Regression with mlr3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Introduction

This vignette demonstrates the `cmprsk.crr` learner from
the `LearnerCompRisksFineGrayCRR` package, which implements
Fine-Gray competing risks regression using `cmprsk::crr`. 
The learner estimates cumulative incidence functions (CIFs)
for competing risks scenarios, supporting fixed and time-varying
ovariates

# Setup

Load required packages and create a task using the `pbc` dataset:

```{r setup}
library(mlr3)
library(mlr3proba)
library(LearnerCompRisksFineGrayCRR)
if (!requireNamespace("cmprsk", quietly = TRUE)) {
  stop("Package 'cmprsk' must be installed")
}
task <- tsk("pbc")
task$set_col_roles(cols = "status", add_to = "stratum")
task$col_roles$feature <- setdiff(task$feature_names, "status")
task$select(c("age", "sex", "bili"))
print(task)
print(table(task$data()[["status"]]))

```

# Training the Learner

Train the Fine-Gray model with time-varying covariates:

```{r train}
learner <- lrn("cmprsk.crr",
  cov2_info = list(
    cov2nms = c("age", "bili"),
    tf = function(uft) cbind(log(uft), log(uft + 1)),
    cov2only = NULL
  ),
  maxiter = 50,
  gtol = 1e-7,
  parallel = FALSE
)
set.seed(123)
part <- partition(task, ratio = 0.7)
learner$train(task, part$train)
print(learner)
```
# Prediction

Predict CIFs for the test set:

```{r predict}
pred <- learner$predict(task, part$test)
print(pred)
```
# Model Diagnostics

Check convergence and variable importance:

```{r diagnostics}
learner$convergence()
learner$importance()
```
References

Fine, J. P., & Gray, R. J. (1999). A Proportional Hazards Model for the Subdistribution of a Competing Risk. Journal of the American Statistical Association, 94(446), 496--509. doi:10.1080/01621459.1999.10474144







